---
title: "All Codes"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulate Data


## Extract Block Maxima Function

```{r}

# function to extract block maxima and plot histogram
Extract_Block_Maxima <- function(num_years, block_size, simulated_SST) {
  num_blocks <- num_years / block_size
  block_size_days <- 365 * block_size
  
  # initialize a numeric vector to store block maxima
  block_maxima <- numeric(num_blocks)
  for (i in 1:num_blocks) {
    # calculate the start and end indices for each block
    start_index <- block_size_days * (i - 1) + 1
    end_index <- block_size_days * i
    block_maxima[i] <- max(simulated_SST[start_index:end_index])
  }
   
  return(block_maxima)
}

```



## GEV-Fitting Function

```{r}

# function to fit GEV model
fit_gev_simulate <- function(block_maxima, block_size) {
  library(ismev)
  gev_fit <- gev.fit(xdat = block_maxima, method = "Nelder-Mead", maxit = 10000)
  return(gev_fit)
}

```



## 10 years

```{r}

# 10 years
set.seed(5623)
num_years <- 10
num_days <- 365 * num_years

# simulated SST data from normal distribution
simulated_SST_10 <- rnorm(365 * 10, mean = 20, sd = 5)

# block size
block_sizes <- c(1, 2, 5)

# create a list to store the fitted results
results_10 <- list()

par(mfrow = c(1, 3))
# loop through block sizes
for (block_size in block_sizes) {
  # extract block maxima
  maxima <- Extract_Block_Maxima(10, block_size, simulated_SST_10)
  # fit GEV model to the block maxima
  results_10[[paste(block_size, "years")]] <- fit_gev_simulate(maxima)
}

# diagnostic plots
par(mfrow = c(1, 3)) 
for (block_size in block_sizes) {
  gev_fit_10 <- results_10[[paste(block_size, "years")]]
  gev.diag(gev_fit_10)
}

```



## 20 years

```{r simulate data}

# 20 years
set.seed(5623)
num_years <- 20
num_days <- 365 * num_years

# simulated SST data from normal distribution
simulated_SST_20 <- rnorm(365 * 20, mean = 20, sd = 5)

# block size
block_sizes <- c(1, 2, 5)
# create a list to store the fitted results
results_20 <- list()

par(mfrow = c(1, 3))
# loop through block sizes
for (block_size in block_sizes) {
  # extract block maxima
  maxima <- Extract_Block_Maxima(20, block_size, simulated_SST_20)
  # fit GEV model to the block maxima
  results_20[[paste(block_size, "years")]] <- fit_gev_simulate(maxima)
}

# diagnostic plots
par(mfrow = c(1, 3)) 
for (block_size in block_sizes) {
  gev_fit_20 <- results_20[[paste(block_size, "years")]]
  gev.diag(gev_fit_20)
}


```



## 50 years

```{r}

# 50 years
set.seed(5623)
num_years <- 50
num_days <- 365 * num_years

# simulated SST data from normal distribution
simulated_SST_50 <- rnorm(365 * 50, mean = 20, sd = 5)

# block size
block_sizes <- c(1, 2, 5)
# create a list to store the fitted results
results_50 <- list()

par(mfrow = c(1, 3))
# loop through block sizes
for (block_size in block_sizes) {
  # extract block maxima
  maxima <- Extract_Block_Maxima(50, block_size, simulated_SST_50)
  # fit GEV model to the block maxima
  results_50[[paste(block_size, "years")]] <- fit_gev_simulate(maxima)
}

# diagnostic plots
par(mfrow = c(1, 3)) 
for (block_size in block_sizes) {
  gev_fit_50 <- results_50[[paste(block_size, "years")]]
  gev.diag(gev_fit_50)
}


```



## 100 years

```{r}

# 100 years
set.seed(5623)
num_years <- 100
num_days <- 365 * num_years

# simulated SST data from normal distribution
simulated_SST_100 <- rnorm(365 * 100, mean = 20, sd = 5)

# block size
block_sizes <- c(1, 2, 5, 10)
# create a list to store the fitted results
results_100 <- list()

par(mfrow = c(1, 3))
# loop through block sizes
for (block_size in block_sizes) {
  # extract block maxima
  maxima <- Extract_Block_Maxima(100, block_size, simulated_SST_100)
  # fit GEV model to the block maxima
  results_100[[paste(block_size, "years")]] <- fit_gev_simulate(maxima)
}

# diagnostic plots
par(mfrow = c(1, 3)) 
for (block_size in block_sizes) {
  gev_fit_100 <- results_100[[paste(block_size, "years")]]
  gev.diag(gev_fit_100)
}


```


## Return Level

### 1yr try

```{r}

library(evd)
library(ismev)

block_maxima_100_1yr <- Extract_Block_Maxima(100, 1, simulated_SST_100)

gev_fit_100_1yr <- fit_gev_simulate(block_maxima_100_1yr)

return_periods <- c(2, 5, 10, 20, 50, 100)

# return levels
gev_return_levels <- evd::qgev(1 - 1/return_periods, loc = gev_fit_100_1yr$mle[1], scale = gev_fit_100_1yr$mle[2], shape = gev_fit_100_1yr$mle[3])

names(gev_return_levels) <- return_periods

print(gev_return_levels)



true_return_levels <- sapply(return_periods, function(T) {
  qnorm((1 - 1/T)^(1/365), mean = 20, sd = 5)
})
names(true_return_levels) <- return_periods

print(true_return_levels)

plot(return_periods, gev_return_levels, type = "b", pch = 19, xlab = "Return Period (years)", ylab = "Return Level (°C)", main = "Return Level Plot", col = "blue")
lines(return_periods, gev_return_levels, col = "blue", lwd = 2)
points(return_periods, true_return_levels, col = "red", pch = 19)
lines(return_periods, true_return_levels, col = "red", lwd = 2)
legend("bottomright", legend = c("GEV Return Levels", "True Normal Quantiles"), col = c("blue", "red"), pch = 19)



comparison <- data.frame(
  Return_Period = return_periods,
  GEV_Return_Level = gev_return_levels,
  True_Return_Level = true_return_levels
)

print(comparison)


```


### 10yr try

```{r}

library(evd)

block_maxima_100_10yr <- Extract_Block_Maxima(100, 10, simulated_SST_100)

gev_fit_100_10yr <- fit_gev_simulate(block_maxima_100_10yr)

return_periods <- c(50, 100, 200, 500)

# return levels
gev_return_levels <- evd::qgev(1 - 1/return_periods, loc = gev_fit_100_10yr$mle[1], scale = gev_fit_100_10yr$mle[2], shape = gev_fit_100_10yr$mle[3])

names(gev_return_levels) <- return_periods

print(gev_return_levels)


true_return_levels <- sapply(return_periods, function(T) {
  qnorm((1 - 1/T)^(1/365), mean = 20, sd = 5)
})
names(true_return_levels) <- return_periods

print(true_return_levels) 

# plot return levels
plot(return_periods, gev_return_levels, type = "b", pch = 19, xlab = "Return Period (years)", ylab = "Return Level (°C)", main = "Return Level Plot", col = "blue", ylim = range(c(gev_return_levels, true_return_levels)))
lines(return_periods, gev_return_levels, col = "blue", lwd = 2)

# add true return levels
points(return_periods, true_return_levels, col = "red", pch = 19)
lines(return_periods, true_return_levels, col = "red", lwd = 2)

# add legend
legend("bottomright", legend = c("GEV Return Levels", "True Normal Quantiles"), col = c("blue", "red"), pch = 19, lwd = 2)


comparison <- data.frame(
  Return_Period = return_periods,
  GEV_Return_Level = gev_return_levels,
  True_Return_Level = true_return_levels
)

print(comparison)

```



### rl function

```{r}

library(evd)


# create a function that generate return levels for each return period for each simulated dataset
simulate_gev_rl <- function(num_years, block_size, return_periods, data) {
  
  # create a matrix to store the return levels for each return period
  sim_rl_list <- matrix(NA, nrow = 100, ncol = length(return_periods))
  colnames(sim_rl_list) <- return_periods
  
  for (i in 1:100) {
    # choose the corresponding dataset
    simulated_SST <- sim_total[[i]]
    # extract block maxima
    block_maxima <- Extract_Block_Maxima(num_years, block_size, simulated_SST)
    # fit GEV model
    gev_fit <- gev.fit(block_maxima, method = "Nelder-Mead", maxit = 10000)
    
    # calculate return levels
    sim_rl <- round(evd::qgev(1 - 1/((return_periods/block_size)), loc = gev_fit$mle[1], scale = gev_fit$mle[2], shape = gev_fit$mle[3]), 2)
    sim_rl_list[i, ] <- sim_rl
  }
  
  return(as.data.frame(sim_rl_list))
}



set.seed(1111)
# create a list of 100 sets of simulated data
sim_total <- lapply(1:100, function(x) rnorm(num_years * 365, mean = 20, sd = 5))


# parameters
num_years <- 100
block_sizes <- c(1, 2, 5, 10)
return_periods <- c(50, 100, 200, 500)
mean <- 20
sd <- 5
data <- sim_total


# true return levels for the normal distribution
true_return_levels <- sapply(return_periods, function(T) {
  qnorm((1 - 1/T)^(1/365), mean = mean, sd = sd)
})
names(true_return_levels) <- return_periods


# return levels for each block size
rl_1 <- simulate_gev_rl(num_years, 1, return_periods, sim_total)
rl_2 <- simulate_gev_rl(num_years, 2, return_periods, sim_total)
rl_5 <- simulate_gev_rl(num_years, 5, return_periods, sim_total)
rl_10 <- simulate_gev_rl(num_years, 10, return_periods, sim_total)


```



### plots


```{r}

# for 1 year block size
par(mfrow = c(2, 2))
# estimated mean return level
mean_rl_1 <- colMeans(rl_1)

# histogram of return levels
for (i in 1:length(return_periods)) {
  hist(rl_1[, i], breaks = 30, main = paste("Return Level for", return_periods[i], "years\nBlock Size: 1 year"), xlab = "Return Level", col = "lightblue", xlim = c(min(rl_1), max(rl_1)))
  abline(v = true_return_levels[i], col = "red", lwd = 2)    # true normal quantile
  abline(v = mean_rl_1[i], col = "blue", lwd = 2, lty = 2)
  legend("topright", legend = c("True Normal Quantile", "Mean Estimated Return Level"), col = c("red","blue"), lwd = 2, lty = c(1, 2), cex = 1, bty = "n", inset = c(-0.35, -0.1))
}

```


```{r}

par(mfrow = c(2, 2))
mean_rl_2 <- colMeans(rl_2)

# for 2-year block size
for (i in 1:length(return_periods)) {
  hist(rl_2[, i], breaks = 30, main = paste("Return Level for", return_periods[i], "years\nBlock Size: 2 years"), xlab = "Return Level", col = "lightblue",xlim = c(min(rl_2), max(rl_2)))
  abline(v = true_return_levels[i], col = "red", lwd = 2)
  abline(v = mean_rl_2[i], col = "blue", lwd = 2, lty = 2)
  legend("topright", legend = c("True Normal Quantile", "Mean Estimated Return Level"), col = c("red","blue"), lwd = 2, lty = c(1, 2), cex = 1, bty = "n", inset = c(-0.35, -0.1))
}

```


```{r}

par(mfrow = c(2, 2))
mean_rl_5 <- colMeans(rl_5)

# for 5-year block size
for (i in 1:length(return_periods)) {
  hist(rl_5[, i], breaks = 30, main = paste("Return Level for", return_periods[i], "years\nBlock Size: 5 years"), xlab = "Return Level", col = "lightblue")
  abline(v = true_return_levels[i], col = "red", lwd = 2)
  abline(v = mean_rl_5[i], col = "blue", lwd = 2, lty = 2)
  legend("topright", legend = c("True Normal Quantile", "Mean Estimated Return Level"), col = c("red","blue"), lwd = 2, lty = c(1, 2), cex = 1, bty = "n", inset = c(-0.35, -0.1))
}


```


```{r}

par(mfrow = c(2, 2))
mean_rl_10 <- colMeans(rl_10)

# for 10-year block size
for (i in 1:length(return_periods)) {
  hist(rl_10[, i], breaks = 30, main = paste("Return Level for", return_periods[i], "years\nBlock Size: 10 years"), xlab = "Return Level", col = "lightblue")
  abline(v = true_return_levels[i], col = "red", lwd = 2)
  abline(v = mean_rl_10[i], col = "blue", lwd = 2, lty = 2)
  legend("topright", legend = c("True Normal Quantile", "Mean Estimated Return Level"), col = c("red","blue"), lwd = 2, lty = c(1, 2), cex = 1, bty = "n", inset = c(-0.4, -0.1))
}

```







# Import Real Data

```{r import data}

library(readxl)
load("Red_Sea_data.RData")
bathymetry <- read_excel("bathymetry.xlsx")
RedSST <- data

```



# EDA

```{r}

library(ggplot2)

# mean SST at each location
mean_by_loc <- colMeans(RedSST)

# organize all variables into one dataframe
mean_SST_loc_df <- data.frame(
  MeanTemp = mean_by_loc,
  Longitude = loc[,1],
  Latitude = loc[,2],
  Depth = bathymetry[,3]
)


# spatical variability
ggplot(mean_SST_loc_df, aes(x = Longitude, y = Latitude, color = MeanTemp)) +
  geom_point(size = 1.5) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Spatial Variability in Sea Surface Temperature (SST)",
       x = "Longitude",
       y = "Latitude",
       color = "Mean SST (°C)")


# mean SST for each day
mean_by_day <- rowMeans(RedSST)

mean_SST_day_df <- data.frame(
  MeanTemp = mean_by_day,
  DayIndex = 1:5840)

ggplot(mean_SST_day_df, aes(x = DayIndex, y = MeanTemp)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Mean Daily Sea Surface Temperature (SST)",
       x = "Day Index",
       y = "Mean SST (°C)") +
  scale_x_continuous(breaks = seq(0, 5840, by = 180)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(mean_SST_day_df[1:730,], aes(x = DayIndex, y = MeanTemp)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Mean Daily Sea Surface Temperature (SST)",
       x = "Day Index",
       y = "Mean SST (°C)") +
  scale_x_continuous(breaks = seq(0, 5840, by = 180)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, hjust = 1))

hist(RedSST, 
     main = "Histogram of Sea Surface Temperature (SST)",
     xlab = "SST (°C)",
     ylab = "Frequency")


```

## Separated

```{r}

library(dplyr)
library(ggplot2)

# mean daily SST in summer and winter
mean_SST_day_df <- rowMeans(RedSST)
summer <- mean_SST_day_df[month %in% c(6,7,8)]
winter <- mean_SST_day_df[month %in% c(12,1,2)]

# line plot of summer mean SST
y_limits <- range(c(max(summer), min(winter)), na.rm = TRUE)
plot(summer, type = "l", col = "red", 
     xlab = "Day", ylab = "Mean SST (°C)",
     main = "Mean Daily Sea Surface Temperature (SST) in Summer and Winter", ylim = y_limits)
# add a line for winter mean SST
lines(winter, type = "l", col = "blue")
legend("topright", legend = c("Summer", "Winter"), col = c("red", "blue"), lty = 1)


# zoom in to the first six months of data
first_six_summer <- summer[1:184]
first_six_winter <- winter[1:180]

plot(first_six_summer, type = "l", col = "red", 
     xlab = "Day", ylab = "Mean SST (°C)",
     main = "Mean Daily Sea Surface Temperature (SST) in Summer and Winter", ylim = y_limits)
# add a line for winter mean SST
lines(first_six_winter, type = "l", col = "blue")
legend("topright", legend = c("Summer", "Winter"), col = c("red", "blue"), lty = 1)

```


# Single Location

## Function Combining BM and GEV Fit

```{r function of find block maxima and GEV fit}

####################################################################################################
# Description:
# Fits a Generalized Extreme Value (GEV) distribution to annual block maxima from daily SST data.
#
# Parameters:
# SST_data: Numeric vector of daily SST values for a single location
#
# Returns:
# GEV fit results
#
# Process:
# 1. Calculate the number of years from SST_data
# 2. Extract annual maxima and store in `block_maxima`.
# 3. Fit GEV distribution to `block_maxima` using `gev.fit`.
# 4. Return GEV fit results.


library(ismev)

fit_gev <- function(SST_data) {
  num_years <- 16
  block_maxima <- numeric(num_years)
  
  # find the maxima for each year
  for (i in 1:num_years) {
    start_index <- 365*i - 364
    end_index <- 365 * i
    block_maxima[i] <- max(SST_data[start_index:end_index])
  }
  
  # fit the GEV distribution to the block maxima (use identity link for all parameters)
  gev_fit <- gev.fit(
    xdat = block_maxima,
    method = "Nelder-Mead",
    maxit = 10000
  )
  
  return(gev_fit)
}

####################################################################################################

```


## 108 MLES 

```{r}

num_locations <- 108

# initialize a list to store MLEs for each location
mles <- matrix(NA, nrow = 108, ncol = 3)

# loop through each location to apply the fit_gev function
for (i in 1:108) {
  RedSST_loc <- RedSST[, i]
  capture.output(gev_fit_single <- fit_gev(RedSST_loc))
  
  # store the mle results
  mles[i, ] <- gev_fit_single$mle
}

# create a dataframe to store the mles for 108 locations
mles_data <- as.data.frame(mles)
colnames(mles_data) <- c("Location", "Scale", "Shape")
print(mles_data)

```



# Incorporate Locations

## Data Alignment 

### Block Maxima

```{r Block Maxima}

# calculate block maxima for each location
block_maxima <- sapply(1:108, function(loc) {
  sapply(1:16, function(year) {
    start_index <- 365*year - 364
    end_index <- year * 365
    max(data[start_index:end_index, loc])
  })
})
  
block_maxima_df <- as.data.frame(block_maxima)
colnames(block_maxima_df) <- paste("Location", 1:108)
rownames(block_maxima_df) <- paste("Maxima for Year", 1:16)
print(block_maxima_df)

```


### Location data

```{r location}

location <- data.frame(loc)

# combine the corresponding location data for each block maxima
location_data <- data.frame(
  longitude = rep(location$lon, each = 16),
  latitude = rep(location$lat, each = 16),
  depth = rep(bathymetry$dep, each = 16),
  log_depth = rep(log(bathymetry$dep), each = 16)
)
print(location_data)


# combine loc and mles
loc_mle <- cbind(loc, mles_data)

# add water depth into loc_mle
loc_mle$dep <- bathymetry$dep

# add log (scale) 
loc_mle$log_dep <- log(loc_mle$dep)

```


### Alignment

```{r fit GEv}

# sequence: Max for year 1 in location 1; max for year 2 in location 1; ... max for year 16 in location 1; max for year 1 in location 2; ... 
block_maxima_vector <- as.vector(t(t(block_maxima_df)))
print(block_maxima_vector)

# location coordinate for location 1 x 16 times; for location 2 x 16 times
location_data <- as.matrix(location_data)
print(location_data)

```



# Location

## ggplots

```{r location scatter plot}

library(ggplot2)

par(mfrow = c(3, 1))

# scatter plot for location parameter with both lon and lat
ggplot(loc_mle, aes(x = lon, y = lat, color = Location)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Location Parameter vs Longitude and Latitude", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Location MLEs") 

# against longitude
ggplot(loc_mle, aes(x = lon, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Location Parameter vs Longitude", x = "Longitude", y = "Location")

# against latitude
ggplot(loc_mle, aes(x = lat, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Location Parameter vs Latitude", x = "Latitude", y = "Location")

# against depth
ggplot(loc_mle, aes(x = dep, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Location Parameter vs Water Depth", x = "Water Depth", y = "Location")

# against log_depth
ggplot(loc_mle, aes(x = log_dep, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Location Parameter vs Log Water Depth", x = "Log Water Depth", y = "Location")

```


Areas with higher longitudes and lower latitudes tend to have higher
average values of extreme temperatures. The relationship tends to be
linear.

Location: 𝜇=𝛽0+𝛽1 Longitude+𝛽2 Latitude 
𝜇=𝛽0+𝛽1 Longitude 
𝜇=𝛽0+𝛽2 Latitude 
find AIC, find the best model (from the negative likelihood ) likelihood ratio test or AIC



## Models


## Ismev   


```{r}

library(ismev)

# location with both lon and lat
gev_fit1 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(1,2),  
  method = "Nelder-Mead",
  maxit = 10000
)


# location with only lon
gev_fit2 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(1),  
  method = "Nelder-Mead",
  maxit = 10000
)


# location with only latitude
gev_fit3 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(2),  
  method = "Nelder-Mead",
  maxit = 10000
)

gev_fit4 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(1,2,4),  
  method = "Nelder-Mead",
  maxit = 10000
)


gev_fit5 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(1,4),  
  method = "Nelder-Mead",
  maxit = 10000
)

gev_fit6 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(2,4),  
  method = "Nelder-Mead",
  maxit = 10000
)

gev_fit7 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(4),  
  method = "Nelder-Mead",
  maxit = 10000
)


```


```{r}

AIC_ismev1 <- 2 * length(gev_fit1$mle) + 2 * gev_fit1$nllh
AIC_ismev2 <- 2 * length(gev_fit2$mle) + 2 * gev_fit2$nllh
AIC_ismev3 <- 2 * length(gev_fit3$mle) + 2 * gev_fit3$nllh
AIC_ismev4 <- 2 * length(gev_fit4$mle) + 2 * gev_fit4$nllh
AIC_ismev5 <- 2 * length(gev_fit5$mle) + 2 * gev_fit5$nllh
AIC_ismev6 <- 2 * length(gev_fit6$mle) + 2 * gev_fit6$nllh
AIC_ismev7 <- 2 * length(gev_fit7$mle) + 2 * gev_fit7$nllh

AIC_ismev <- data.frame(
  Model = c("ismev_fit1", "ismev_fit2", "ismev_fit3", "ismev_fit4", "ismev_fit5", "ismev_fit6", "ismev_fit7"),
  AIC = c(AIC_ismev1, AIC_ismev2, AIC_ismev3, AIC_ismev4, AIC_ismev5, AIC_ismev6, AIC_ismev7)
)
print(AIC_ismev)

```

The fourth model has the smallest AIC. 
𝜇=𝛽0+𝛽1 Longitude+𝛽2 Latitude + 𝛽3 log_water_depth






# Scale

## ggplots

```{r scatter plot for scale parameter}

library(ggplot2)
par(mfrow = c(3, 1))
# add log (scale) 
loc_mle$log_Scale <- log(loc_mle$Scale)

ggplot(loc_mle, aes(x = lon, y = lat, color = Scale)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Scale Parameter vs Longitude and Latitude", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Scale") 

ggplot(loc_mle, aes(x = lon, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Log Scale Parameter vs Longitude", x = "Longitude", y = "Log Scale")


ggplot(loc_mle, aes(x = lat, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Latitude", y = "Log Scale")

```

more purple near the coast, more red inside (maybe distance from the
coast/ or water depth)
maybe the darker purple point in the middle is an island
(##?? The point in the middle has 789m depth, which is not an island, but
the scale of its temperature is very small. )



## Bathymetry

```{r water depth}

# plot of water depth at each location
ggplot(bathymetry, aes(x = lon, y = lat, color = dep)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Water Depth at Each Location", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Depth (m)")

# plot of the relationship of scale parameter and the water depth
ggplot(loc_mle, aes(x = dep, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Log Scale Parameter vs Water Depth", x = "Water Depth", y = "Log Scale")

# plot of the relationship of scale parameter and the log water depth
ggplot(loc_mle, aes(x = log_dep, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Log Scale Parameter vs Log Water Depth", x = "Log Water Depth", y = "Log Scale")

```

The relationship appears weak but positive.






## Texmex

```{r}

library(texmex)

# combine block maxima with the location data and water depth
location_data_df <- as.data.frame(cbind(BlockMaxima = block_maxima_vector, location_data))


# create gev models using texmex

# scale with both lon and lat
evm_fit1 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude + latitude)
print(evm_fit1)

# scale with only lon
evm_fit2 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude)
print(evm_fit2)

# scale with only lat
evm_fit3 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ latitude)
print(evm_fit3)

# scale with lon, lat and log_depth
evm_fit4 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude + latitude + log_depth)
print(evm_fit4)


evm_fit5 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude + log_depth)
print(evm_fit5)

evm_fit6 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ latitude + log_depth)
print(evm_fit6)


evm_fit7 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ log_depth)
print(evm_fit7)


```


```{r}

AIC_evm1 <- AIC(evm_fit1)
AIC_evm2 <- AIC(evm_fit2)
AIC_evm3 <- AIC(evm_fit3)
AIC_evm4 <- AIC(evm_fit4)
AIC_evm5 <- AIC(evm_fit5)
AIC_evm6 <- AIC(evm_fit6)
AIC_evm7 <- AIC(evm_fit7)

# create a data frame to store the AIC values
AIC_evm <- data.frame(
  Model = c("evm_fit1", "evm_fit2", "evm_fit3", "evm_fit4", "evm_fit5", "evm_fit6", "evm_fit7"),
  AIC = c(AIC_evm1, AIC_evm2, AIC_evm3, AIC_evm4, AIC_evm5, AIC_evm6, AIC_evm7)
)

print(AIC_evm)


```



## All AIC


```{r}

AIC_results_combined <- rbind(AIC_ismev, AIC_evm)

# Order the combined results from smallest to largest AIC value
AIC_results_combined <- AIC_results_combined[order(AIC_results_combined$AIC), ]

# Print the combined and ordered AIC results
print(AIC_results_combined)

```

evm_fit5 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude + log_depth)
print(evm_fit3)




# Shape

## ggplots

```{r shape paramter}

par(mfrow = c(3, 1))
ggplot(loc_mle, aes(x = lon, y = lat, color = Shape)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Shape Parameter vs Longitude and Latitude", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Shape") 

ggplot(loc_mle, aes(x = lon, y = Shape)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + # Linear trend line
  labs(title = "Shape Parameter vs Longitude", x = "Longitude", y = "Shape")


ggplot(loc_mle, aes(x = lat, y = Shape)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + # Linear trend line
  labs(title = "Shape Parameter vs Latitude", x = "Latitude", y = "Shape")


```

in environmental analysis, normally assume shape to be identical, can
just fixed



# Final Model

## final model1

```{r}

gev_final1 <- (evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ longitude + latitude + log_depth,
    phi = ~ longitude + log_depth))
print(gev_final1)

```

μ=31.93+0.09627⋅longitude−0.2116⋅latitude+0.0515⋅log(depth)
log(σ)=0.8539−0.02388⋅longitude−0.02996⋅log(depth)
ξ=−0.3779

might be multicolinearity : NaN in SE



## multicolinearity

```{r}

# define the coordinates of the reference location
location1 <- c(34.75,28.75)
location_matrix <- as.matrix(location)

# compute distances from the reference to all other locations
distances <- dist(rbind(location1, location_matrix))

distances_matrix <- as.matrix(distances)

# extract distances from reference
distances_to_L1 <- distances_matrix[1, -1] 

# add the column of distances to the dataframe
location_data <- as.data.frame(location_data)
location_data$distance <- rep(distances_to_L1, each = 16)
location_data$log_distance <- log(location_data$distance)
print(location_data)

```

```{r}

ggplot(location_data, aes(x = longitude, y = latitude, color = distance)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Distance from each Coordinate to the Reference Point", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Distance") 

ggplot(location_data, aes(x = longitude, y = latitude, color = log_distance)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Log Distance from each Coordinate to the Reference Point", 
       x = "Longitude", 
       y = "Latitude", 
       color = "Log Distance") 

```



```{r}

loc_mle$log_distance <- log(distances_to_L1)
loc_mle$distance <- distances_to_L1

ggplot(loc_mle, aes(x = distance, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Distance", y = "Log Scale")

ggplot(loc_mle, aes(x = log_distance, y = log_Scale)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Log_Distance", y = "Log Scale")

ggplot(loc_mle, aes(x = distance, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Distance", y = "Location")

ggplot(loc_mle, aes(x = log_distance, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Log_Distance", y = "Location")

ggplot(loc_mle, aes(x = dep, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Depth", y = "Location")

ggplot(loc_mle, aes(x = log_dep, y = Location)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Log Scale Parameter vs Latitude", x = "Log_depth", y = "Location")

```





## ismev

```{r}

library(ismev)
location_data <- as.matrix(location_data)
# location: with both lon and lat
gev_fit8 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(5),  
  method = "Nelder-Mead",
  maxit = 10000
)


gev_fit9 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(3,5),  
  method = "Nelder-Mead",
  maxit = 10000
)


gev_fit10 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(3),  
  method = "Nelder-Mead",
  maxit = 10000
)

gev_fit11 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(4),  
  method = "Nelder-Mead",
  maxit = 10000
)

gev_fit12 <- gev.fit(
  xdat = block_maxima_vector,
  ydat = location_data,
  mul = c(4,6),  
  method = "Nelder-Mead",
  maxit = 10000
)


```


```{r}

AIC_ismev8 <- 2 * length(gev_fit8$mle) + 2 * gev_fit8$nllh
AIC_ismev9 <- 2 * length(gev_fit9$mle) + 2 * gev_fit9$nllh
AIC_ismev10 <- 2 * length(gev_fit10$mle) + 2 * gev_fit10$nllh
AIC_ismev11 <- 2 * length(gev_fit11$mle) + 2 * gev_fit11$nllh
AIC_ismev12 <- 2 * length(gev_fit12$mle) + 2 * gev_fit12$nllh


AIC_ismev_new <- data.frame(
  Model = c("ismev_fit8", "ismev_fit9", "ismev_fit10", "ismev_fit11", "ismev_fit12"),
  AIC = c(AIC_ismev8, AIC_ismev9, AIC_ismev10, AIC_ismev11, AIC_ismev12))
print(AIC_ismev_new)

```


## texmex

```{r}

library(texmex)

# combine block maxima with the location data and water depth
location_data_df <- as.data.frame(cbind(BlockMaxima = block_maxima_vector, location_data))


# create gev models using texmex
evm_fit8 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance + log_depth,
    phi = ~ log_distance + log_depth)
print(evm_fit8)


evm_fit9 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance + log_depth,
    phi = ~ log_distance)
print(evm_fit9)


evm_fit10 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance + log_depth,
    phi = ~ log_depth)
print(evm_fit10)


evm_fit11 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance,
    phi = ~ log_distance + log_depth)
print(evm_fit11)


evm_fit12 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance,
    phi = ~ log_distance)
print(evm_fit12)


evm_fit13 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance,
    phi = ~ log_depth)
print(evm_fit13)


evm_fit14 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_depth,
    phi = ~ log_distance + log_depth)
print(evm_fit14)


evm_fit15 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_depth,
    phi = ~ log_distance)
print(evm_fit15)


evm_fit16 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_depth,
    phi = ~ log_depth)
print(evm_fit16)



```

```{r}

AIC_evm8 <- AIC(evm_fit8)
AIC_evm9 <- AIC(evm_fit9)
AIC_evm10 <- AIC(evm_fit10)
AIC_evm11 <- AIC(evm_fit11)
AIC_evm12 <- AIC(evm_fit12)
AIC_evm13 <- AIC(evm_fit13)
AIC_evm14 <- AIC(evm_fit14)
AIC_evm15 <- AIC(evm_fit15)
AIC_evm16 <- AIC(evm_fit16)
# create a data frame to store the AIC values
AIC_evm_new <- data.frame(
  Model = c("evm_fit8", "evm_fit9", "evm_fit10","evm_fit11", "evm_fit12", "evm_fit13","evm_fit14", "evm_fit15", "evm_fit16"),
  AIC = c(AIC_evm8, AIC_evm9, AIC_evm10, AIC_evm11, AIC_evm12, AIC_evm13, AIC_evm14, AIC_evm15, AIC_evm16)
)

print(AIC_evm_new)

```



## final model 2

```{r}

library(texmex)
location_data_df <- as.data.frame(cbind(BlockMaxima = block_maxima_vector, location_data))

final_model2 <- evm(BlockMaxima, data=location_data_df, family=gev, 
    mu = ~ log_distance,
    phi = ~ log_distance + log_depth)
print(final_model2)

```



# Diagnostics

## Quantile Plot


```{r}

library(ggplot2)

# mean SST at each location
mean_by_loc <- colMeans(RedSST)

# specify the row numbers to highlight
highlight_rows <- c(1, 22, 44, 70, 90, 108)

# Create a data frame for plotting
mean_SST_loc_df <- data.frame(
  MeanTemp = mean_by_loc,
  Longitude = location[,1],
  Latitude = location[,2],
  Depth = bathymetry[,3],
  Highlight = ifelse(seq_len(nrow(location)) %in% highlight_rows, "Highlight", "Normal")
)

# Plotting spatial variability
ggplot(mean_SST_loc_df, aes(x = Longitude, y = Latitude, color = MeanTemp)) +
  geom_point(size = 1.5) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Spatial Variability in Sea Surface Temperature (SST)",
       x = "Longitude",
       y = "Latitude",
       color = "Mean SST (°C)") +
  geom_point(data = mean_SST_loc_df[highlight_rows, ], 
             aes(x = Longitude, y = Latitude), 
             color = "black", 
             size = 3, 
             shape = 21, 
             fill = "yellow") +  # Customize the highlight color and shape
  theme_minimal()


```

### loc1


```{r}

sigma <- exp(0.6909 + (-0.2856 * loc_mle$log_distance) + (-0.07912 * loc_mle$log_dep))
xi <- -0.366
mu <- 28.11 + (1.711 * loc_mle$log_distance)

# Coefficient estimates for location 1
sigma_loc1 <- sigma[1]  
xi_loc1 <- -0.366
mu_loc1 <- mu[1]        

library(evd)  

# the probability for each quantile
probability_values <- 1:16 / 17
# sort block maxima for location 1
sorted_bm_loc1 <- sort(block_maxima_df[,1]) 

# empirical quantile
empirical_quantiles_loc1 <- quantile(sorted_bm_loc1, probs = probability_values)

# theoretical quantiles using GEV parameters
theoretical_quantiles_loc1 <- qgev(probability_values, mu_loc1, sigma_loc1, xi_loc1)

# define the axis range
axis_limits <- range(c(theoretical_quantiles_loc1, empirical_quantiles_loc1))

# Q-Q plot
plot(theoretical_quantiles_loc1, empirical_quantiles_loc1,
     main = "Q-Q Plot for Location 1",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,  
     xlim = axis_limits, 
     ylim = axis_limits) 
abline(0, 1, col = "blue")  # diagonal line for reference


```



### All

```{r}


par(mfrow = c(2, 3))

# Coefficient estimates directly from your provided data
sigma_loc1 <- sigma[1]  # assuming 'sigma' vector is defined in your workspace
xi_loc1 <- -0.366
mu_loc1 <- mu[1]        # assuming 'mu' vector is defined in your workspace

library(evd)  # Assuming 'evd' package is used for qgev

# Assume sigma_loc1, xi_loc1, and mu_loc1 are already calculated or provided

# Calculate the probabilities
probability_values <- 1:16 / 17


sorted_bm_loc1 <- sort(block_maxima_df[,1]) 
# Assuming 'sorted_bm_loc1' is your sorted block maxima data for location 1
# You need to have this data sorted and loaded into your environment
empirical_quantiles_loc1 <- quantile(sorted_bm_loc1, probs = probability_values)

# Calculate the theoretical quantiles using GEV parameters
theoretical_quantiles_loc1 <- qgev(probability_values, mu_loc1, sigma_loc1, xi_loc1)

# Define the axis limits based on the quantiles
axis_limits <- range(c(theoretical_quantiles_loc1, empirical_quantiles_loc1))

# Plotting the Q-Q plot
plot(theoretical_quantiles_loc1, empirical_quantiles_loc1,
     main = "Q-Q Plot for Location 1",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,  # Using a solid circle for better visibility
     xlim = axis_limits, # Set x-axis limits
     ylim = axis_limits) # Set y-axis limits
abline(0, 1, col = "blue")  # diagonal line for reference






# Coefficient estimates directly from your provided data for Location 22
sigma_loc22 <- sigma[22]  # assuming 'sigma' vector is defined in your workspace
xi_loc22 <- -0.366        # same xi as before, assuming it remains constant across locations
mu_loc22 <- mu[22]        # assuming 'mu' vector is defined in your workspace

# Calculate the probabilities
probability_values <- 1:16 / 17

sorted_bm_loc22 <- sort(block_maxima_df[,22]) 
# Assuming 'sorted_bm_loc22' is your sorted block maxima data for location 22
empirical_quantiles_loc22 <- quantile(sorted_bm_loc22, probs = probability_values)

# Calculate the theoretical quantiles using GEV parameters
theoretical_quantiles_loc22 <- qgev(probability_values, mu_loc22, sigma_loc22, xi_loc22)

# Define the axis limits based on the quantiles
axis_limits <- range(c(theoretical_quantiles_loc22, empirical_quantiles_loc22))

# Plotting the Q-Q plot for Location 22
plot(theoretical_quantiles_loc22, empirical_quantiles_loc22,
     main = "Q-Q Plot for Location 22",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,  # Using a solid circle for better visibility
     xlim = axis_limits, # Set x-axis limits
     ylim = axis_limits) # Set y-axis limits
abline(0, 1, col = "blue")  # diagonal line for reference




# Coefficient estimates for Location 44
sigma_loc44 <- sigma[44]  # 'sigma' vector must be defined in your workspace
xi_loc44 <- -0.366        # Assuming xi is consistent across locations
mu_loc44 <- mu[44]        # 'mu' vector must be defined in your workspace

# Calculate probabilities for the quantiles
probability_values <- 1:16 / 17

# Sorted block maxima data for location 44
sorted_bm_loc44 <- sort(block_maxima_df[,44]) 
# Calculate empirical quantiles
empirical_quantiles_loc44 <- quantile(sorted_bm_loc44, probs = probability_values)

# Calculate theoretical quantiles using the GEV model
theoretical_quantiles_loc44 <- qgev(probability_values, mu_loc44, sigma_loc44, xi_loc44)

# Define the axis limits based on the quantiles
axis_limits <- range(c(theoretical_quantiles_loc44, empirical_quantiles_loc44))

# Plotting the Q-Q plot for Location 44
plot(theoretical_quantiles_loc44, empirical_quantiles_loc44,
     main = "Q-Q Plot for Location 44",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,  # Using a solid circle for better visibility
     xlim = axis_limits, # Set x-axis limits
     ylim = axis_limits) # Set y-axis limits
abline(0, 1, col = "blue")  # diagonal line for reference





sigma_loc70 <- sigma[70]  # 'sigma' vector must be defined in your workspace
xi_loc70 <- -0.366        # Assuming xi is consistent across locations
mu_loc70 <- mu[70]        # 'mu' vector must be defined in your workspace

# Calculate probabilities for the quantiles
probability_values <- 1:16 / 17

# Sorted block maxima data for location 70
sorted_bm_loc70 <- sort(block_maxima_df[,70]) 
# Calculate empirical quantiles
empirical_quantiles_loc70 <- quantile(sorted_bm_loc70, probs = probability_values)

# Calculate theoretical quantiles using the GEV model
theoretical_quantiles_loc70 <- qgev(probability_values, mu_loc70, sigma_loc70, xi_loc70)

# Define the axis limits based on the quantiles
axis_limits <- range(c(theoretical_quantiles_loc70, empirical_quantiles_loc70))

# Plotting the Q-Q plot for Location 70
plot(theoretical_quantiles_loc70, empirical_quantiles_loc70,
     main = "Q-Q Plot for Location 70",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,
     xlim = axis_limits,
     ylim = axis_limits)
abline(0, 1, col = "blue")  # diagonal line for reference




sigma_loc90 <- sigma[90]  # 'sigma' vector must be defined in your workspace
xi_loc90 <- -0.366        # Assuming xi is consistent across locations
mu_loc90 <- mu[90]        # 'mu' vector must be defined in your workspace

# Sorted block maxima data for location 90
sorted_bm_loc90 <- sort(block_maxima_df[,90]) 
# Calculate empirical quantiles
empirical_quantiles_loc90 <- quantile(sorted_bm_loc90, probs = probability_values)

# Calculate theoretical quantiles using the GEV model
theoretical_quantiles_loc90 <- qgev(probability_values, mu_loc90, sigma_loc90, xi_loc90)

# Plotting the Q-Q plot for Location 90
plot(theoretical_quantiles_loc90, empirical_quantiles_loc90,
     main = "Q-Q Plot for Location 90",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,
     xlim = axis_limits,
     ylim = axis_limits)
abline(0, 1, col = "blue")




sigma_loc108 <- sigma[108]  # 'sigma' vector must be defined in your workspace
xi_loc108 <- -0.366        # Assuming xi is consistent across locations
mu_loc108 <- mu[108]

# Calculate the probabilities
probability_values <- 1:16 / 17

sorted_bm_loc108 <- sort(block_maxima_df[,108]) 

# Empirical quantiles for location 108
empirical_quantiles_loc108 <- quantile(sorted_bm_loc108, probs = probability_values)

# Theoretical quantiles for location 108
# Note: Make sure mu_loc108, phi_loc108, and xi_loc108 are calculated correctly beforehand
theoretical_quantiles_loc108 <- qgev(probability_values, mu_loc108, sigma_loc108, xi_loc108)

# Define the axis limits based on the quantiles
axis_limits <- range(c(theoretical_quantiles_loc108, empirical_quantiles_loc108))

# Plotting the Q-Q plot
plot(theoretical_quantiles_loc108, empirical_quantiles_loc108,
     main = "Q-Q Plot for Location 108",
     xlab = "Theoretical Quantiles (Model)",
     ylab = "Empirical Quantiles",
     pch = 1,  # Using a solid circle for better visibility
     xlim = axis_limits, # Set x-axis limits
     ylim = axis_limits) # Set y-axis limits
abline(0, 1, col = "blue")  # diagonal line for reference, changed to red for better visibility










```






## Density plot


### All

```{r}

par(mfrow = c(2, 3))

# generate a sequence of 100 equally spaced values 
x_vals_loc1 <- seq(min(sorted_bm_loc1), max(sorted_bm_loc1), length.out = 100)
# GEV density for location 1
gev_density_loc1 <- dgev(x_vals_loc1, mu_loc1, sigma_loc1, xi_loc1)

# empirical density histogram
hist(sorted_bm_loc1, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 1", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc1, gev_density_loc1, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc1), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)




# GEV density for location 22
x_vals_loc22 <- seq(min(sorted_bm_loc22), max(sorted_bm_loc22), length.out = 100)
gev_density_loc22 <- dgev(x_vals_loc22, mu_loc22, sigma_loc22, xi_loc22)

# empirical density histogram
hist(sorted_bm_loc22, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 22", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc22, gev_density_loc22, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc22), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)




# GEV density for location 44
x_vals_loc44 <- seq(min(sorted_bm_loc44), max(sorted_bm_loc44), length.out = 100)
gev_density_loc44 <- dgev(x_vals_loc44, mu_loc44, sigma_loc44, xi_loc44)

# empirical density histogram
hist(sorted_bm_loc44, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 44", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc44, gev_density_loc44, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc44), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)





# GEV density for location 70
x_vals_loc70 <- seq(min(sorted_bm_loc70), max(sorted_bm_loc70), length.out = 100)
gev_density_loc70 <- dgev(x_vals_loc70, mu_loc70, sigma_loc70, xi_loc70)

# empirical density histogram
hist(sorted_bm_loc70, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 70", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc70, gev_density_loc70, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc70), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)





# GEV density for location 90
x_vals_loc90 <- seq(min(sorted_bm_loc90), max(sorted_bm_loc90), length.out = 100)
gev_density_loc90 <- dgev(x_vals_loc90, mu_loc90, sigma_loc90, xi_loc90)

# empirical density histogram
hist(sorted_bm_loc90, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 90", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc90, gev_density_loc90, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc90), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)




# GEV density for location 108
x_vals_loc108 <- seq(min(sorted_bm_loc108), max(sorted_bm_loc108), length.out = 100)
gev_density_loc108 <- dgev(x_vals_loc108, mu_loc108, sigma_loc108, xi_loc108)

# empirical density histogram
hist(sorted_bm_loc108, 
     probability = TRUE,  
     breaks = 10,         
     main = "Density Plot for Location 108", 
     xlab = "Block Maxima", 
     ylab = "Density",
     col = "lightgray",   
     border = "black")    

# theoretical density curve
lines(x_vals_loc108, gev_density_loc108, col = "blue", lwd = 2)

# empirical density curve
lines(density(sorted_bm_loc108), col = "black", lwd = 2)

# legend to the plot
legend("topright", legend = c("Empirical", "Model"), col = c("black", "blue"), lwd = 2)

```




# Return Level

## function

```{r}

# function for calculating return levels for 1 year block size
gev_rl <- function(return_period) {
  
  # calculate the probability corresponding to the return period
  p <- 1 - 1 / return_period
  
  # the estimated coefficients
  sigma <- exp(0.6909 + (-0.2856 * loc_mle$log_distance) + (-0.07912 * loc_mle$log_dep))
  xi <- -0.366
  mu <- 28.11 + (1.711 * loc_mle$log_distance)
  
  # calculate the corresponding quantile of the gev distribution
  return_level <- evd::qgev(p, mu, sigma, xi)
  
  return(return_level)
}


```


## rls

```{r}

return_periods <- c(10, 20, 50, 100, 200, 500)

# initialize a return matrix
return_levels <- matrix(NA, nrow = length(return_periods), ncol = nrow(loc_mle))

# calculate return levels for each return periods
for (i in seq_along(return_periods)) {
  return_levels[i, ] <- round(gev_rl(return_periods[i]),3)
}

# set rownames as the return period and columnnames as location
rownames(return_levels) <- paste(return_periods, "years")
colnames(return_levels) <- paste("Location", 1:nrow(loc_mle))

return_levels <- as.data.frame(return_levels)
print(return_levels)

print(return_levels[,c(1,22,44,70,90,108)])

```



## All

```{r}

par(mfrow = c(2, 3))

return_periods_continuous <- seq(10, 500, by = 1)

# calculate return levels for the continuous return periods
return_levels_1 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu[1], sigma[1], xi)
})

# calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# calculate standard errors (SE) for scale and add confidence intervals
se_scale <- z_value * sigma[1] * 0.1  # avoid non-positive values

lower_bound_1 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu[1], pmax(0.0001, sigma[1] - se_scale), xi)  # avoid negative values
})

upper_bound_1 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu[1], sigma[1] + se_scale, xi)
})

# calculate the empirical return levels using block maxima
empirical_return_levels_1 <- quantile(block_maxima_df[,1], probs = 1 - 1/return_periods_continuous)

# set y-axis limits to ensure all points are visible
y_limits <- range(c(return_levels_1, empirical_return_levels_1, lower_bound_1, upper_bound_1))

# Return level plot
plot(return_periods, return_levels_1, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 1", ylim = y_limits)

# add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_1, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_1, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_1, col = "blue", pch = 19, cex = 0.5)

legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)





## Location 22

# Calculate return levels for the specified return periods
return_levels_22 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu_loc22, sigma_loc22, xi_loc22)
})

# Calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# Calculate standard errors (SE) for scale and add confidence intervals
se_scale_22 <- z_value * sigma_loc22 * 0.1  # Adjust as necessary to avoid non-positive values

lower_bound_22 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc22, pmax(0.0001, sigma_loc22 - se_scale_22), xi_loc22)
})

upper_bound_22 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc22, sigma_loc22 + se_scale_22, xi_loc22)
})

# Calculate the empirical return levels using block maxima
empirical_return_levels_22 <- quantile(block_maxima_df[, 22], probs = 1 - 1/return_periods_continuous)

# Set y-axis limits to ensure all points are visible
y_limits_22 <- range(c(return_levels_22, empirical_return_levels_22, lower_bound_22, upper_bound_22))

# Return level plot for Location 22
plot(return_periods, return_levels_22, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 22", ylim = y_limits_22)

# Add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_22, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_22, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_22, col = "blue", pch = 19, cex = 0.5)

# Add a legend to the plot
legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)




# location 44
# Calculate return levels for the specified return periods
return_levels_44 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu_loc44, sigma_loc44, xi_loc44)
})

# Calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# Calculate standard errors (SE) for scale and add confidence intervals
se_scale_44 <- z_value * sigma_loc44 * 0.1  # Adjust as necessary to avoid non-positive values

lower_bound_44 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc44, pmax(0.0001, sigma_loc44 - se_scale_44), xi_loc44)
})

upper_bound_44 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc44, sigma_loc44 + se_scale_44, xi_loc44)
})

# Calculate the empirical return levels using block maxima
empirical_return_levels_44 <- quantile(block_maxima_df[, 44], probs = 1 - 1/return_periods_continuous)

# Set y-axis limits to ensure all points are visible
y_limits_44 <- range(c(return_levels_44, empirical_return_levels_44, lower_bound_44, upper_bound_44))

# Return level plot for Location 44
plot(return_periods, return_levels_44, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 44", ylim = y_limits_44)

# Add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_44, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_44, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_44, col = "blue", pch = 19, cex = 0.5)

# Add a legend to the plot
legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)



# location 70 

# Calculate return levels for the specified return periods
return_levels_70 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu_loc70, sigma_loc70, xi_loc70)
})

# Calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# Calculate standard errors (SE) for scale and add confidence intervals
se_scale_70 <- z_value * sigma_loc70 * 0.1  # Adjust as necessary to avoid non-positive values

lower_bound_70 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc70, pmax(0.0001, sigma_loc70 - se_scale_70), xi_loc70)
})

upper_bound_70 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc70, sigma_loc70 + se_scale_70, xi_loc70)
})

# Calculate the empirical return levels using block maxima
empirical_return_levels_70 <- quantile(block_maxima_df[, 70], probs = 1 - 1/return_periods_continuous)

# Set y-axis limits to ensure all points are visible
y_limits_70 <- range(c(return_levels_70, empirical_return_levels_70, lower_bound_70, upper_bound_70))

# Return level plot for Location 70
plot(return_periods, return_levels_70, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 70", ylim = y_limits_70)

# Add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_70, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_70, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_70, col = "blue", pch = 19, cex = 0.5)

# Add a legend to the plot
legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)



# location 90
return_levels_90 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu_loc90, sigma_loc90, xi_loc90)
})

# Calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# Calculate standard errors (SE) for scale and add confidence intervals
se_scale_90 <- z_value * sigma_loc90 * 0.1  # Adjust as necessary to avoid non-positive values

lower_bound_90 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc90, pmax(0.0001, sigma_loc90 - se_scale_90), xi_loc90)
})

upper_bound_90 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc90, sigma_loc90 + se_scale_90, xi_loc90)
})

# Calculate the empirical return levels using block maxima
empirical_return_levels_90 <- quantile(block_maxima_df[, 90], probs = 1 - 1/return_periods_continuous)

# Set y-axis limits to ensure all points are visible
y_limits_90 <- range(c(return_levels_90, empirical_return_levels_90, lower_bound_90, upper_bound_90))

# Return level plot for Location 90
plot(return_periods, return_levels_90, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 90", ylim = y_limits_90)

# Add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_90, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_90, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_90, col = "blue", pch = 19, cex = 0.5)

# Add a legend to the plot
legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)
      



# location 108
      
return_levels_108 <- sapply(return_periods, function(period) {
  p <- 1 - 1 / period
  evd::qgev(p, mu_loc108, sigma_loc108, xi_loc108)
})

# Calculate confidence intervals for return levels
alpha <- 0.05  # 95% confidence interval
z_value <- qnorm(1 - alpha / 2)  # Z-value for the confidence interval

# Calculate standard errors (SE) for scale and add confidence intervals
se_scale_108 <- z_value * sigma_loc108 * 0.1  # Adjust as necessary to avoid non-positive values

lower_bound_108 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc108, pmax(0.0001, sigma_loc108 - se_scale_108), xi_loc108)
})

upper_bound_108 <- sapply(return_periods_continuous, function(period) {
  p <- 1 - 1 / period
  qgev(p, mu_loc108, sigma_loc108 + se_scale_108, xi_loc108)
})

# Calculate the empirical return levels using block maxima
empirical_return_levels_108 <- quantile(block_maxima_df[, 108], probs = 1 - 1/return_periods_continuous)

# Set y-axis limits to ensure all points are visible
y_limits_108 <- range(c(return_levels_108, empirical_return_levels_108, lower_bound_108, upper_bound_108))

# Return level plot for Location 108
plot(return_periods, return_levels_108, type = "l", col = "black", lwd = 2,
     xlab = "Return Period (Years)", ylab = "Return Level",
     main = "Location 108", ylim = y_limits_108)

# Add confidence interval dashed lines
lines(return_periods_continuous, lower_bound_108, col = "red", lty = 2)
lines(return_periods_continuous, upper_bound_108, col = "red", lty = 2)
points(return_periods_continuous, empirical_return_levels_108, col = "blue", pch = 19, cex = 0.5)

# Add a legend to the plot
legend("topright", 
       legend = c("Return Level", "95% CI", "Empirical"), 
       col = c("black", "red", "blue"), 
       lty = c(1, 2, NA), 
       lwd = c(2, 1, NA), 
       pch = c(NA, NA, 19), 
       cex = 0.55)
      
     

```






# Multivariate


```{r}

# Assuming summer_SST is a matrix where columns represent different locations
summer_SST <- RedSST[month %in% c(6, 7, 8), ]
summer_SST <- as.matrix(summer_SST)

# Extract SST data for location 1
sst_location1 <- summer_SST[, 1]

# List of locations to compare with location 1
locations_to_compare <- c(2, 22, 44, 70, 90, 108)

# Set up plotting area for multiple plots
par(mfrow = c(2, 3))  # Adjust the layout as needed

# Loop through the list of locations and create scatter plots
for (i in locations_to_compare) {
  # Extract SST data for the current location
  sst_current_location <- summer_SST[, i]
  
  # Create the scatter plot
  plot(sst_location1, sst_current_location,
       xlab = "SST at Location 1",
       ylab = paste("SST at Location", i),
       main = paste("SST at Location 1 vs. Location", i),
       col = "blue",
       pch = 19,  # Point type
       cex = 0.7)  # Point size
}


```






```{r}

# 23 has highest 1755m depth, 25 has only 4m depth but they are very close
par(mfrow = c(1, 3))

sst_location23 <- summer_SST[, 23]
sst_location25 <- summer_SST[, 25]
  plot(sst_location23, sst_location25,
       xlab = "SST at Location 23",
       ylab = paste("SST at Location 25"),
       main = paste("SST at Location 23 vs. Location 25"),
       col = "blue",
       pch = 19,  # Point type
       cex = 0.7)  # Point size
  
  
# 66 has depth of 2m
sst_location23 <- summer_SST[, 23]
sst_location66 <- summer_SST[, 66]
  plot(sst_location23, sst_location66,
       xlab = "SST at Location 23",
       ylab = paste("SST at Location 66"),
       main = paste("SST at Location 23 vs. Location 66"),
       col = "blue",
       pch = 19,  # Point type
       cex = 0.7)  # Point size
  
# 104 has depth of 2m
sst_location23 <- summer_SST[, 23]
sst_location104 <- summer_SST[, 104]
  plot(sst_location23, sst_location104,
       xlab = "SST at Location 23",
       ylab = paste("SST at Location 104"),
       main = paste("SST at Location 23 vs. Location 104"),
       col = "blue",
       pch = 19,  # Point type
       cex = 0.7)  # Point size
  
  
  
```




# End











